---
title: "Final Exam"
output:
  html_document: default
  pdf_document: default
date: "2022-12-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Final Exam
Data Retrieved from Kaggle, but originally from FiveThirtyEight
https://www.kaggle.com/datasets/thedevastator/uncovering-insights-to-college-majors-and-their?select=all-ages.csv

This data set provides information on college majors and their employability, popularity, and salary. 

#Inputting and Setting up Data/Packages
z   Q7 
```{r}
mydata <- read.csv("C:\\Users\\bucol\\Desktop\\recent-grads.csv")
library(tidyverse)  # data manipulation
library(factoextra) # clustering algorithms & visualization
library(ISLR)
library(cluster)
library(flexclust)

#First I will modifying the data to turn the variables of interest into rates that are all positive in direction. We will be turning everything into a rate instead of normalizing. 

newdata <- mydata
newdata$Employmen_rate <- mydata$Employed / mydata$Total # This is the postive version of unemployment rate
newdata$College_Jobs_Rate <- mydata$College_jobs / mydata$Total #This is the rate of college jobs
newdata$NonLow_Wage_jobs_Rate <- (mydata$Total - mydata$Low_wage_jobs) / mydata$Total #This is the rate of non low jwage 
newdata$Median <- mydata$Median / 110000 #this turns the median salary into a rate where we're using the highest salary as the divisor

#Then I will be removing several variables due to redundancy. 
#to do this I will create a list of variables. 

myvars <- c("Employmen_rate","Median","College_Jobs_Rate", "NonLow_Wage_jobs_Rate") 
#This is the selection of variables I will use to provide some basic information about recent graduates. 

majordata <- newdata[,myvars] #This is now selecting only those variables. 
summary(majordata) 


```

For this analysis we will consider the variables importance as seen below:

Highest: College Job Rate (This shows the rate of grads who got a job that uses their college level education)
Medium: Non low wage jobs and Median 
Lowest: Employment rate

Due to this, we will weight each value as shown below:

College Jobs Rate: 3
Median: 1.5
NonLow Jobs Rate: 1.5
Employment Rate: 1

Instead of normalizing, we have created rates for each value. Now we will weight them. 

```{r}
weighteddata <- majordata
weighteddata$Median <- majordata$Median *1.5
weighteddata$NonLow_Wage_jobs_Rate <- majordata$NonLow_Wage_jobs_Rate *1.5
weighteddata$College_Jobs_Rate <- majordata$College_Jobs_Rate *3

distancedata <- get_dist(weighteddata)
fviz_dist(distancedata)
fviz_nbclust(weighteddata, kmeans, method = "wss") 
#It seems our elbow point is 2, but its sort of hard to tell. 
fviz_nbclust(weighteddata, kmeans, method = "silhouette")
#This method suggests 2 clusters.
gap_stat <- clusGap(weighteddata, FUN = kmeans, nstart = 25,
                    K.max = 10, B = 50)
fviz_gap_stat(gap_stat)
#this method suggests 6 clusters. 
```
#Now we cluster. 

```{r}
set.seed(17)
cluster2 <- kmeans(weighteddata, centers = 2, nstart = 25)

#Let's Visualize
cluster2$centers
fviz_cluster(cluster2, data= weighteddata) 
cluster2$size


#Visualizing this, we can see we have two cluster, but theres not a ton of information was can draw from this since the groups are so large. Lets break it up a bit more like the gap stat suggested. 

cluster6 <- kmeans(weighteddata, centers = 6, nstart = 25)

#Let's Visualize
cluster6$centers
fviz_cluster(cluster6, data= weighteddata) 
cluster6$size

```
Here we can see the first cluster has two groups. 
Group 1 has 66 majors
Group 2 has 107. 

Here we can see the second cluster has six groups. 
Group 1 has 21 majors
Group 2 has 16
Group 3 has 56 
Group 4 has 14
Group 5 has 31
Group 6 has 35

Due to the nature of the problem, I will be utilizing the 6 clusters for the analysis. 

```{r}
cluster6dataA <- cluster6$cluster
cluster6dataB <- cbind(newdata, cluster6dataA) #this adds our kmeans clusters for 6 clusters to the dataset
c2vars <- c("Major","Major_category", "Total","ShareWomen", "Employmen_rate","Median","College_Jobs_Rate", "NonLow_Wage_jobs_Rate","cluster6dataA")
cluster6data <- cluster6dataB[,c2vars]

Group1 <- subset(cluster6data, cluster6dataA ==1) #this returns just the majors in the 1st cluster
Group2 <- subset(cluster6data, cluster6dataA == 2) #this returns just the majors in the 2nd cluster
Group3 <- subset(cluster6data, cluster6dataA == 3) #this returns just the majors in the 3rd cluster
Group4 <- subset(cluster6data, cluster6dataA == 4) #this returns just the majors in the 4th cluster
Group5 <- subset(cluster6data, cluster6dataA == 5) #this returns just the majors in the 5th cluster
Group6 <- subset(cluster6data, cluster6dataA == 6) #this returns just the majors in the 6th cluster

Group6
```






