---
title: "Assignment_5"
output:
  html_document: default
  pdf_document: default
  word_document: default
date: "2022-11-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Inputting and Setting up Data/Packages

```{r}
mydata <- read.csv("C:\\Users\\bucol\\Downloads\\Cereals.csv")
library(tidyverse)  # data manipulation
library(factoextra) # clustering algorithms & visualization
library(cluster)
library(caret)
set.seed(49)


summary(mydata)
```
#We see there is some missing data that needs to be removed
```{r}
cerealscomplete <- na.omit(mydata)
cerealscomplete #this has removed the cereals with incomplete data

#Now we need to make the cereal names the row names
row.names(cerealscomplete) <- cerealscomplete[,1]
#now we must remove the names from the dataset
cerealscomplete <- cerealscomplete[,-1]

#I'm going to remove the first two columns because they're categorical. 
cerealscomplete <- cerealscomplete[,-1]
cerealscomplete <- cerealscomplete[,-1]
cerealscomplete

```

#Since we are dealing with very different variables, we must normalize the data
```{r}
normcereal <- scale(cerealscomplete)
summary(normcereal) #now our data is scaled.
```

#Now we will use Agnes to compare single, complete, and average linkage and Ward. 

```{r}
set.seed(17)
hc_single <- agnes(normcereal, method = "single")
hc_complete <- agnes(normcereal, method = "complete")
hc_average <- agnes(normcereal, method = "average")
hc_ward <- agnes(normcereal, method = "ward")

#The we compare agglomerative coefficients
print(hc_single$ac) #0.6067859
print(hc_complete$ac) #0.8353712
print(hc_average$ac) #0.7766075
print(hc_ward$ac) #0.9046042  (this one is the highest)

```

#Based on the agglomerative coefficients, the ward method is the best. 
#we will now plot the dendrogram
```{r}
plot_ward <- pltree(hc_ward, cex = 0.6, hang = -1)
plot_ward
```
#Looking at the dendrogram, I think 6 clusters seems pretty good. 
```{r}
pltree(hc_ward, cex = 0.6, hang = -1)
rect.hclust(hc_ward,k=6,border = 1:6)
                               
```
#Now we check their stability. 
#First, we partition. 
```{r}
set.seed(17)

#Creating partition set of 50%
partset <- createDataPartition(normcereal[,"calories"], p=0.5, list = FALSE)
CerA <- normcereal[partset,] #this is the first 50%
CerB <- normcereal[-partset,]#second set

#Now we'll run the AGNES to both sets. 
a_ward <-agnes(CerA, method = "ward")
print(a_ward$ac) #0.8498682
b_ward <-agnes(CerB, method = "ward")
print(b_ward$ac) #0.8383771
#we can note that the agglomerative coefficients are very similar. 

plot_a <- pltree(a_ward, cex=0.6, hang = -1)
plot_b <- pltree(b_ward, cex=0.6, hang = -1)
```
If I compare these partitioned dendrograms to the original, we can see that most cereals remain in the same cluster. 
This tells us that it is pretty stable. 

#Senario: Find cluster of healthy cereal. 
Okay, since we want healthy cereals we should be considering the nutrional items in this data set and their importance. 
We probably are looking at some that should be high and some that would be better low. 
high: protein, fiber, potassium, vitamins, and honestly rating since we want the kids to actually eat it.  
low: fat, sodium
now calories and carbs are also important, but I dont think we specifically want them high or low since it needs to be
balanced between healthy and filling. 
For the purpose of this analysis, I would to fet rid of shelf.  
To be honest, looking at this data, its a bit unbalanced because of the serving sizes which I assume is what cups is. 
If we really wanted to consider these on an equal playing field, we should probably make a ratio of the nutritional aspect
per its cups so that all the nutritional data is scaled similarly. ie, a cereal doesn't look healthier because its just less
food. 
Also, due to the importance of scale here, you would want to manipulate your data depending on health benefit importance, 
so normalizing the data would not be the first option since we do care about the actual values of the nutritional facts. 



```

```


