---
output:
  pdf_document: default
  html_document: default
---
#Assignment 1, getting to know R ##Data set "Impact of Beauty on Instructor's Teaching Ratings" from <https://vincentarelbundock.github.io/Rdatasets/articles/data.html>

**Inputting Data**

```{r}
 mydata <- read.csv("C:\\Users\\bucol\\Downloads\\TeachingRatings.csv")
print(mydata)
```

Now I can call on this data throughout the rest of my document!

**Descriptive Statistics-quantitative**

```{r}
summary(mydata$age) 
```

The standard deviation for age was `r sd(mydata$age)`. The summary feature provides lots of descriptive statistics, but it doesn't include standard deviation, which is why I have added that with "inline R coding"as opposed to a block. 

```{r}
hist(mydata$age)
```

```{r}
summary(mydata$beauty) 

```

The standard deviation for age was `r sd(mydata$beauty)`

```{r}
hist(mydata$beauty)
```

```{r}
summary(mydata$eval)
```

The standard deviation for age was `r sd(mydata$eval)`

```{r}
hist(mydata$eval)
```

**Descriptive Statistics-qualitative**

Since these are categorical, we're looking mostly for frequencies. 

```{r}
g<-table(mydata$gender)
g
gpercent<-prop.table(g)*100
gpercent

```

```{r}
div<-table(mydata$division)
div
divpercent<-prop.table(div)*100
divpercent

```

Now trying this with the package Hmisc. First, I have to call that package with the library function, then I can use the "describe" feature. 

```{r}
library(Hmisc)
describe(mydata$minority)
```

```{r}
describe(mydata$tenure)
```

That was definitely simpler. :)

**Plots!**

When I do this, I chose two variables to plot, one on the x-axis, one on the y-axis. Then I label them so with their actualy variabel titles not their program titles. 
```{r}
plot(mydata$beauty,mydata$eval, xlab="Beauty", ylab = "Evaluation", main = "Beauty vs Evaluation")
```

Here's some boxplots!
```{r}
boxplot(mydata$age)
```

```{r}
boxplot(mydata$eval)
```



**Transforming with log**

We'll start with the variable "students"

```{r}
hist(mydata$students)
```

This histogram is a bit skewed, so we can transform the data into its logs to analyze:

```{r}
transform<-log10(mydata$students)
transform
hist(transform)
```

Now the data can be analyzed. Yay R!
