---
title: "Assignment_4"
output:
  word_document: default
  html_document: default
date: "2022-10-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Inputting and Setting up Data/Packages
**A**
```{r}
mydata <- read.csv("C:\\Users\\bucol\\Downloads\\Pharmaceuticals.csv")
library(tidyverse)  # data manipulation
library(factoextra) # clustering algorithms & visualization
library(ISLR)
set.seed(49)

PharmData <- select(mydata, 3:11)
summary(PharmData)

```
Now we need to normalize our data since Euclidean distance is very affected by scale. Then we will get our distances. 

```{r}
NormPharm <- scale(PharmData)
distance <- get_dist(NormPharm)
fviz_dist(distance)
fviz_nbclust(NormPharm, kmeans, method = "wss") 
#It's a bit challenging to read this graph to determine the elbow point, so we'll also check witht the silhouette method
fviz_nbclust(NormPharm, kmeans, method = "silhouette")
#From here, we know to choose 5 clusters. 

```
#Now we cluster!
```{r}
set.seed(49)
clusterk5 <- kmeans(NormPharm, center = 5, nstart = 25)

#Let's Visualize
clusterk5$centers
fviz_cluster(clusterk5, data= NormPharm)
clusterk5$size
```
**b**
Looking at our cluster plot, we can see:

Cluster 1: rows 6, 8, 12
Cluster 2: rows 5, 9, 14, 20
Cluster 3: rows 11, 13, 15, 17
Cluster 4: rows 1, 3, 4, 7, 10, 16, 19, 21
Cluster 5: rows 2, 18

**c**
Comparing this the variables not used in the cluster analysis, there aren't very discernable patterns. 
-The countries seem pretty randomly distributed throughout the clusters.  
-The only two non NYSE companies are in cluster 1.
-The reccommendations also seem relatively random. 

**d**
Naming Clusters

Cluster 1: Lowest Market Capitalization, Return on Equity, and Asset Turnover, but highest Beta and Return on Assets
Cluster 2: Lowest Price/earnings ratio, but highest Revenue Growth
Cluster 3: Lowest Leverage, but highest Return on equity, return on assets, asset turnover, net profit margin and market capitalization
Cluster 4: Lowest revenue growth
Cluster 5: Lowest beta, but highest Price/earnings ratio 





